Windowed least-squares model reduction for dynamical systems

### 2019-10-07 main.tex
in progress

\address[a]{Sandia National Laboratories,  Livermore, CA}
\begin{abstract}
This work proposes a \methodNameLower\ (\methodAcronym) model-reduction
	formulation for dynamical systems. The proposed approach sequentially minimizes the
	time-continuous full-order-model residual over a low-dimensional space--time trial
	subspace defined over an arbitrarily defined time window. The approach comprises a generalization 
of existing model reduction approaches, as particular instances of
  the methodology associate with Galerkin,
	least-squares Petrov--Galerkin (LSPG), and space–time LSPG projection. In
	addition, the approach
	addresses key deficiencies in existing model-reduction
	techniques, e.g., the dependence of (space--time) LSPG projection on the
	time discretization and the exponential growth in time exhibited by \textit{a posteriori}
	error bounds for Galerkin and LSPG projection.  We consider two types of
	space--time trial
	subspaces within the proposed framework: one that reduces only the spatial
	dimension of the full-order model, and one that reduces both the spatial and
	temporal dimensions of the full-order model.  For each type of trial
	subspace, we consider two different solution techniques: direct (i.e.,
	discretize-then-optimize) and indirect (i.e., optimize-then-discretize).
	Numerical experiments conducted using the trial subspaces characterized by spatial
	dimension reduction demonstrate that the \methodAcronym\
	approach can yield more accurate solutions that better satisfy the governing equations than
	Galerkin and LSPG projection.

\begin{comment}
This work proposes a \methodNameLower\ (\methodAcronym) model-reduction
	formulation for dynamical systems.  
The proposed approach minimizes the
	time-continuous full-order-model residual over a low-dimensional 
	trial subspace across a sequence of time windows. % that decompose the time
We consider two types of trial subspaces
	within this framework:
	one that reduces the spatial dimension of the full-order model, and one that
	reduces both the spatial and temporal dimensions of the full-order model. 
We consider two methods to solve the residual-minimization problems: direct
	(i.e., discretize-then-optimize) and indirect (i.e.,
	optimize-then-discretize).
The approach addresses key deficiencies in a variety 
of existing model-reduction approaches; e.g., the time-step and time-scheme dependence of least-squares Petrov--Galerkin (LSPG), the exponentially growing error bounds present in Galerkin and LSPG, and the offline cost of space--time LSPG. 
Various limiting cases of the proposed method yield existing reduced-order modeling techniques, namely Galerkin, Least-Squares Petrov–Galerkin, and Space–Time Least-Squares Petrov– Galerkin.
Numerical experiments on the compressible Euler and Navier--Stokes equations demonstrate that, at an increased computational cost, the \methodAcronym\ formulation can yield more accurate and physically relevant solutions than the Galerkin and Least-Squares Petrov--Galerkin approaches. 
\end{comment}
\end{abstract}
\end{frontmatter}

@@

Simulating parameterized dynamical systems arises in many applications across
science and engineering, including design, control, uncertainty
quantification, and epidemiology, for example. In many contexts, executing a dynamical-system
simulation at a single parameter instance---which entails the numerical
integration of a system of ordinary differential equations (ODEs)---incurs an
extremely large computational cost.  This occurs, for example, when the
state-space dimension is large (e.g., due to fine spatial resolution when
discretizing a partial differential equation) and/or when the number of time
instances is large (e.g., due to time-step limitations incurred by stability
or accuracy considerations).  When the application is time critical or many
query in nature, analysts must replace such large-scale parameterized
dynamical-system models (which we refer to as the full-order model) with a low-cost approximation that makes the
application tractable.

Projection-based reduced-order models (ROMs) comprise one such approximation
strategy. First, these techniques execute a computationally expensive
\textit{offline} stage that computes a low-dimensional \textit{trial subspace} on
which the dynamical-system state can be well approximated (e.g., by computing
state ``snapshots'' at different time and parameter instances, by solving
Lyapunov equations). Second, these methods execute an inexpensive
\textit{online} stage during which they compute approximations to the
dynamical-system trajectory that reside on this trial subspace via a
projection process.

Model reduction for linear-time-invariant systems (and other well structured
dynamical systems) is quite mature
\cite{wilcox_benner_rev,moore,roberts,GugercinIRKA}, as system-theoretic properties (e.g., controllability,
observability, asymptotic stability, $\mathcal H_2$-optimality) can be readily
quantified and accounted for in the projection process; this often results in
reduced-order models that inherit such important properties.  The primary
challenge in
developing reduced-order models for general nonlinear dynamical systems is
that such properties are difficult to assess quantitatively. As a result, it
is difficult to develop reduced-order models that preserve important
dynamical-system properties, which often results in methods that yield
trajectories that are
inaccurate, unstable, or violate physical properties.  To address this, researchers have pursued several directions that
aim to imbue reduced-order models for nonlinear
dynamical systems with properties that can improve robustness and accuracy.
These
efforts include residual-minimizing
projection that equips the ROM solution with a notion of optimality~\cite{carlberg_lspg,carlberg_gnat,legresley_1,legresley_2,legresley_3,bui_resmin_steady,bui_unsteady,rovas_thesis,carlberg_thesis,bui_thesis,l1};
space--time
projection that leads to error bounds that grow slowly in time~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC,benner_st};
``energy-based'' inner
products that ensure non-increasing entropy in the ROM
solution~\cite{rowley_pod_energyproj,Kalashnikova_sand2014,chan2019entropy};
basis-adaptation methods that improve the ROM's accuracy \textit{a
posteriori}~\cite{carlberg_hadaptation,adeim_peherstorfer,etter2019online}, stabilizing subspace rotations that account for truncated modes \textit{a priori}~\cite{basis_rotation}, structure-preserving
methods that enforce conservation \cite{carlberg_conservative_rom} or
(port-)Hamiltonian/Lagrangian structure
\cite{LALL2003304,carlberg2012spd,structurePreserveBeattie,chaturantabut2016structure,farhat2014dimensional} in the ROM; and subgrid-scale
modeling methods that aim to improve accuracy by addressing the closure
problem~\cite{san_iliescu_geostrophic,iliescu_pod_eddyviscosity,iliescu_vms_pod_ns,Bergmann_pod_vms,iliescu_ciazzo_residual_rom,Wang_ROM_thesis,wentland_apg,Wang:269133,San2018}.
We note that these techniques are often not mutually exclusive.
%Residual-minimizing projection methods are the most relevant to the current work and comprise the focus of the following review.
Residual-minimizing and space--time projection are the most relevant classes of
methods for the current work and comprise the focus of the following review.

Residual-minimization methods in model reduction compute the solution within a
low-dimensional trial subspace that minimizes the full-order-model residual.
Researchers have developed such residual-minimizing model-reduction methods
for both static systems (i.e., systems without time-dependence)
\cite{legresley_1,legresley_2,legresley_3,bui_resmin_steady,rovas_thesis,carlberg_thesis,bui_thesis}
and dynamical
systems~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg,carlberg_lspg_v_galerkin}.
In the latter category,
Refs.~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg}
formulated the residual minimization problem for dynamical systems by
sequentially minimizing the \textit{time-discrete} full-order-model residual
(i.e., the residual arising after applying time discretization) at each time
instance on the time-discretization grid. This formulation is often referred
to as the \textit{least--squares Petrov--Galerkin} (LSPG) method.
Ref.~\cite{carlberg_lspg_v_galerkin} performed detailed analyses of this
formulation and examined its connections with Galerkin projection. Critically,
this work demonstrated that (1) under certain conditions, LSPG can be derived
from a Petrov--Galerkin projection executed on the time-continuous
full-order-model residual, and (2) LSPG and Galerkin projection are equivalent
in the limit as the time step goes to zero (i.e., Galerkin projection
minimizes the time-instantaneous full-order-model residual). Numerous
numerical experiments have demonstrated that LSPG projection often yields more
accurate and stable solutions than Galerkin
projection~\cite{bui_thesis,carlberg_lspg_v_galerkin,carlberg_gnat,carlberg_thesis,parish_apg}.
The common intuitive explanation for this improved performance is that, by
minimizing the full-order-model residual over a finite time window (rather
than time instantaneously), LSPG projection computes solutions that are more
accurate over a larger part of the trajectory as compared to Galerkin projection.

However, the LSPG approach has several notable shortcomings. First, LSPG
performance exhibits a complex dependence on the time discretization. In
particular, changing
the time step modifies both the
time window over which LSPG minimizes the residual as well as the
time-discretization error of the full-order model on which LSPG is based. 
In particular, because LSPG and Galerkin projection are equivalent in the
limit as the time step goes to zero, LSPG accuracy approaches the (poor)
accuracy of Galerkin as the time step shrinks, but this also has the effect of
reducing the time-dicretization error of the full-order model.
As a
consequence, LSPG often yields the smallest error for an intermediate value of the
time step (see, e.g., Ref.~\cite[Figure 9]{carlberg_lspg_v_galerkin}); there
is no known way to compute this optimal time step \textit{a priori}.
Second, as the LSPG approach performs sequential residual minimization in
time, its \textit{a posteriori} error bounds grow exponentially in time
\cite{carlberg_lspg_v_galerkin}, and it is not equipped with any notion of
optimality over the entire time domain of interest. As a result, LSPG is not
equipped with \textit{a priori} guarantees of accuracy or stability, even for
linear time-invariant systems~\cite{bui_thesis}.

Researchers have pursued the development of space--time
residual-minimization
projection~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC} to
address the issues incurred by sequential residual minimization in time.
Space--time residual-minimization approaches compute the space--time
trajectory within a low-dimensional space--time trial subspace (spanned by
a small number of space--time basis vectors) that minimizes the full-order-model
residual over the entire time domain.
We note that space--time residual minimization yields a fundamentlaly
different model than sequential residual minimization in time. In particular, space--time
residual minimization associates with an algebraic system of equations defined
over all space and time, whose solution comprises a \text{fixed} vector of
(space--time) generalized coordinate; on the other hand, sequential residual
minimization in time (often) associates with a system of ODEs whose solution
comprises a \textit{time-dependent} vector of (spatial) generalized
coordinates.
As a result, space--time residual minimization yields models that are equipped with \textit{a priori} error bounds that grow more slowly
in time, and the resulting \textit{trajectory} exhibits a notion of
optimality; this implies that the solution exhibits a notion of optimality
over the entire time domain. However, these techniques also suffer from several
limitations. First, the computational cost of solving the algebraic system arising from space--time
projection scales cubically with the number of space--time degrees of freedom;
in contrast, the computational cost incurred by standard
spatial-projection-based ROMs is linear in the number of temporal degrees of
freedom, as the attendant solvers can leverage the lower-block triangular
structure of the system arising from the sequential nature of time
evolution. As a result, solving the algebraic systems arising from space--time
projection is generally intractable without applying hyper-reduction in time
\cite{choi_stlspg,constantine_strom}. Second, space--time residual
minimization precludes future
state prediction, as these methods employ space--time basis vectors defined over the
entire time domain of interest, which must have been including in the training
simulations.

The objectives of this work are to overcome the shortcomings of existing
residual-minimizing methods, and to provide a unifying framework from which
existing methods can be assessed. In essence, the proposed  \textit{\methodNameLower\
(\methodAcronym)} method sequentially minimizes the FOM
residual over a sequence of arbitrarily defined
time windows. The method is characterized by three notable aspects.  First, the method
minimizes the \textit{time-continuous}  residual (i.e., that associated with
the full-order model ODE). By adopting a time-continuous viewpoint, the
formulation decouples the underlying temporal discretization scheme from the
residual-minimization problem, thus addressing a key deficiency of both LSPG
and space--time LSPG projection. Critically, time-continuous residual minimization also exposes
two different solution methods:  a \textit{discretize-then-optimize}
(i.e., direct) method, and an \textit{optimize-then-discretize} (i.e.,
indirect) method. Second, the method sequentially minimizes
the residual over arbitrarily defined \textit{time windows} rather than
sequentially minimizing the residual over time steps (as in LSPG projection)
or over the entire time domain (as in space--time residual-minimization
methods). This equips the method with additionally flexibility that enables it
to explore more fine-grained tradeoffs between computational cost and error.
Finally, \methodAcronym\ is formulated for two kinds of space--time trial
subspaces: one that reduces only the spatial dimension of the full-order
model (as employed by traditional spatial-projection-based methods), and one that reduces both the spatial and temporal dimensions of the
full-order model (as employed by space--time residual-minimization methods).
The above attributes allow that \methodAcronym\ method to be viewed as a
generalization of existing model-reduction methods, as 
Galerkin, LSPG, and space--time LSPG projection correspond to specific
instances of the formulation.
Figure~\ref{fig:flowchart} depicts how the proposed \methodAcronym\ method
provides a unifying framework from which these existing approaches can be derived.

The \methodAcronym\ approach can be viewed as a hybrid space--time method and displays 
commonalities with several related efforts. 
First, Ref.~\cite{bui_thesis} tersely formulated a space--time least-squares ROM
and connected this formulation with optimal control; specifically it mentioned the
optimize-then-discretize vs.\ discretize-then-optimize approaches. This work
did not fully develop this approach, eschewing it 
for 
sequential residual minimization in time (i.e., LSPG). The present work thus
formally develops and extends several of the concepts put forth
in Ref.~\cite{bui_thesis}. Next, Ref.~\cite{constantine_strom} developed a
space--time residual minimization formulation for model interpolation.  The
present work distinguishes itself from Ref.~\cite{constantine_strom} in that
(1) this work also considers trial subspaces characterized by a reduction of
the spatial dimension of the FOM only, and (2) we
minimize the time-continuous FOM residual over arbitrary time
windows. We note that,
similar to the current work, Ref.~\cite{constantine_strom} employs
minimization of the time-continuous FOM residual as a starting point; as such,
this work shares some thematic similarities with
Ref.~\cite{constantine_strom}.  Lastly, Ref.~\cite{choi_stlspg} develops a
space--time extension of the LSPG approach that minimizes the time-discrete
FOM residual over the entire time domain.  The present work distinguishes
itself from Ref.~\cite{choi_stlspg} in that (1) this work minimizes the
time-continuous FOM residual, (2) this work minimizes this residual over
arbitrary time windows, and (3) this work considers trial subspaces
characterized by a reduction of the spatial dimension of the FOM only.

@@

 In summary, specific contributions of this work include:
\begin{enumerate}
%\item The development of a time-continuous residual minimization framework for reduced-order models.
\item The \methodNameLower\ (\methodAcronym) framework for dynamical-system
	model reduction. The approach sequentially minimizes the time-continuous
		full-order model residual over arbitrary time windows.
\item Support of two space--time trial subspaces: one that reduces the spatial
	dimension of the FOM only, and one that reduces both the spatial and
		temporal dimensions of the FOM. The former case is of particular
		interest in the \methodAcronym\  contex, as the stationary conditions are
		derived via the Euler--Lagrange equations and comprise a coupled two-point
		Hamiltonian boundary value problem containing a forward and backward
		system. The forward system, which is forced by an auxiliary costate,
		evolves the (spatial) generalized coordinates of the ROM in time. The
		backward system, which is forced by the time-continuous FOM residual
		evaluated about the ROM state, governs the dynamics of the costate. 
	\item Derivation of two solution techniques:
		\textit{discretize-then-optimize} (i.e., direct) and
		\textit{optimize-then-discretize} (i.e., indirect). 
	\item Derivations of conditions under which the
		\methodAcronym\ framework recovers Galerkin, LSPG, and space--time LSPG
		projection.
\item Error analysis of the \methodAcronym\ framework using
	trial subspaces characterized by a reduction of
the spatial dimension of the FOM only.
	This analysis demonstrates that the \methodAcronymROMs\ error 
		is bounded \textit{a priori} by a combination of the
		error at the start of the window and the integrated FOM ODE residual
		evaluated at the FOM state projected onto the trial subspace. As a result,
		the error bound exhibits exponential growth in the number of time windows
		(not total time instances), yielding smaller error-bound growth in thime than
		Galerkin or LSPG projection. \KTC{Check this}
\item Numerical experiments for 
	trial subspaces characterized by a reduction of
the spatial dimension of the FOM only, which demonstrate two key findings:
\begin{itemize}
\item Minimizing the residual over a larger time window leads to more stable
	and physically relevant solutions \KTC{What is `physically relevant'?} as
		well as smaller space--time residual norms. 
\item Minimizing the residual over a larger time window does not necessarily
	lead to a more accurate trajectory (as measured in the space--time
		$\elltwo$-norm of the solution). Instead, minimizing the residual over an
		intermediate-sized time window leads to the smallest trajectory error.
\end{itemize}
\end{enumerate}

>>>>

\address[a]{Sandia National Laboratories,  Livermore, CA}
\begin{abstract}
This work proposes a \methodNameLower\ (\methodAcronym) model-reduction
	formulation for dynamical systems. The proposed approach sequentially minimizes the
	time-continuous full-order-model residual over a low-dimensional space--time trial
	subspace defined over an arbitrarily defined time window. The approach comprises a generalization 
of existing model reduction approaches, as particular instances of
  the methodology recover the Galerkin,
	least-squares Petrov--Galerkin (LSPG), and space–time LSPG approaches. In
	addition, the approach
	addresses key deficiencies in existing model-reduction
	techniques, e.g., the dependence of (space--time) LSPG on the
	time discretization and the exponential growth in time exhibited by \textit{a posteriori}
	error bounds for both Galerkin and LSPG.  We consider two types of
	space--time trial
	subspaces within the proposed framework: one that reduces only the spatial
	dimension of the full-order model, and one that reduces both the spatial and
	temporal dimensions of the full-order model.  For each type of trial
	subspace, we consider two different solution techniques: direct (i.e.,
	discretize-then-optimize) and indirect (i.e., optimize-then-discretize).
	Numerical experiments conducted using the trial subspaces characterized by spatial
	dimension reduction demonstrate that the \methodAcronym\
	approach can yield more accurate solutions that better satisfy the governing equations than
	the Galerkin and LSPG approaches.

\end{abstract}
\end{frontmatter}

@@

Simulating parameterized dynamical systems arises in many applications across
science and engineering. In many contexts, executing a dynamical-system
simulation at a single parameter instance---which entails the numerical
integration of a system of ordinary differential equations (ODEs)---incurs an
extremely large computational cost.  This occurs, for example, when the
state-space dimension is large (e.g., due to fine spatial resolution when
discretizing a partial differential equation) and/or when the number of time
instances is large (e.g., due to time-step limitations incurred by stability
or accuracy considerations).  When the application is time critical or many
query in nature, analysts must replace such large-scale parameterized
dynamical-system models (which we refer to as the full-order model) with a low-cost approximation that makes the
application tractable.

Projection-based reduced-order models (ROMs) comprise one such approximation
strategy. First, these techniques execute a computationally expensive
\textit{offline} stage that computes a low-dimensional \textit{trial subspace} on
which the dynamical-system state can be well approximated (e.g., by computing
state ``snapshots'' at different time and parameter instances, by solving
Lyapunov equations). Second, these methods execute an inexpensive
\textit{online} stage during which they compute approximations to the
dynamical-system trajectory that reside on this trial subspace via, e.g., 
residual minimization.

Model reduction for linear-time-invariant systems (and other well structured
dynamical systems) is quite mature
\cite{wilcox_benner_rev,moore,roberts,GugercinIRKA}, as system-theoretic properties (e.g., controllability,
observability, asymptotic stability, $\mathcal H_2$-optimality) can be readily
quantified and accounted for; this often results in
reduced-order models that inherit such important properties.  The primary
challenge in
developing reduced-order models for general nonlinear dynamical systems is
that such properties are difficult to assess quantitatively. As a result, it
is difficult to develop reduced-order models that preserve important
dynamical-system properties, which often results in methods that yield
trajectories that are
inaccurate, unstable, or violate physical properties.  To address this, researchers have pursued several directions that
aim to imbue reduced-order models for nonlinear
dynamical systems with properties that can improve robustness and accuracy.
These
efforts include residual-minimization approaches that equip the ROM solution with a notion of optimality~\cite{carlberg_lspg,carlberg_gnat,legresley_1,legresley_2,legresley_3,bui_resmin_steady,bui_unsteady,rovas_thesis,carlberg_thesis,bui_thesis,l1};
space--time approaches that leads to error bounds that grow slowly in time~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC,benner_st};
``energy-based'' inner
products that ensure non-increasing entropy in the ROM
solution~\cite{rowley_pod_energyproj,Kalashnikova_sand2014,chan2019entropy};
basis-adaptation methods that improve the ROM's accuracy \textit{a
posteriori}~\cite{carlberg_hadaptation,adeim_peherstorfer,etter2019online}, stabilizing subspace rotations that account for truncated modes \textit{a priori}~\cite{basis_rotation}, structure-preserving
methods that enforce conservation \cite{carlberg_conservative_rom} or
(port-)Hamiltonian/Lagrangian structure
\cite{LALL2003304,carlberg2012spd,structurePreserveBeattie,chaturantabut2016structure,farhat2014dimensional} in the ROM; and subgrid-scale
modeling methods that aim to improve accuracy by addressing the closure
problem~\cite{san_iliescu_geostrophic,iliescu_pod_eddyviscosity,iliescu_vms_pod_ns,Bergmann_pod_vms,iliescu_ciazzo_residual_rom,Wang_ROM_thesis,wentland_apg,Wang:269133,San2018}.
We note that these techniques are often not mutually exclusive.
%Residual-minimizing projection methods are the most relevant to the current work and comprise the focus of the following review.
Residual-minimizing and space--time approaches are the most relevant classes of
methods for the current work and comprise the focus of the following review.

Residual-minimization methods in model reduction compute the solution within a
low-dimensional trial subspace that minimizes the full-order-model residual.
Researchers have developed such residual-minimizing model-reduction methods
for both static systems (i.e., systems without time-dependence)
\cite{legresley_1,legresley_2,legresley_3,bui_resmin_steady,rovas_thesis,carlberg_thesis,bui_thesis}
and dynamical
systems~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg,carlberg_lspg_v_galerkin}.
In the latter category,
Refs.~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg}
formulated the residual minimization problem for dynamical systems by
sequentially minimizing the \textit{time-discrete} full-order-model residual
(i.e., the residual arising after applying time discretization) at each time
instance on the time-discretization grid. This formulation is often referred
to as the \textit{least--squares Petrov--Galerkin} (LSPG) method.
Ref.~\cite{carlberg_lspg_v_galerkin} performed detailed analyses of this
formulation and examined its connections with the Galerkin approach. Critically,
this work demonstrated that (1) under certain conditions, LSPG can be cast as a 
Petrov--Galerkin approach applied to the time-continuous
full-order-model residual, and (2) the LSPG and Galerkin approaches are equivalent
in the limit as the time step goes to zero (i.e., the Galerkin approach 
minimizes the time-instantaneous full-order-model residual). Numerous
numerical experiments have demonstrated that LSPG often yields more
accurate and stable solutions than Galerkin
~\cite{bui_thesis,carlberg_lspg_v_galerkin,carlberg_gnat,carlberg_thesis,parish_apg}.
The common intuitive explanation for this improved performance is that, by
minimizing the full-order-model residual over a finite time window (rather
than time instantaneously), LSPG computes solutions that are more
accurate over a larger part of the trajectory as compared to Galerkin.

However, LSPG has several notable shortcomings. First, LSPG
exhibits a complex dependence on the time discretization. In
particular, changing
the time step modifies both the
time window over which LSPG minimizes the residual as well as the
time-discretization error of the full-order model on which LSPG is based. 
In particular, because LSPG and Galerkin are equivalent in the
limit as the time step goes to zero, the accuracy of LSPG approaches the (poor)
accuracy of Galerkin as the time step shrinks, but this also has the effect of
reducing the time-dicretization error of the full-order model.
As a
consequence, LSPG often yields the smallest error for an intermediate value of the
time step (see, e.g., Ref.~\cite[Figure 9]{carlberg_lspg_v_galerkin}); there
is no known way to compute this optimal time step \textit{a priori}.
Second, as the LSPG approach performs sequential residual minimization in
time, its \textit{a posteriori} error bounds grow exponentially in time
\cite{carlberg_lspg_v_galerkin}, and it is not equipped with any notion of
optimality over the entire time domain of interest. As a result, LSPG is not
equipped with \textit{a priori} guarantees of accuracy or stability, even for
linear time-invariant systems~\cite{bui_thesis}.

Researchers have pursued the development of space--time
residual-minimization approaches~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC} to
address the issues incurred by sequential residual minimization in time.
Space--time residual-minimization approaches compute the space--time
trajectory within a low-dimensional space--time trial subspace (spanned by
a small number of space--time basis vectors) that minimizes the full-order-model
residual over the entire time domain.
We note that space--time residual minimization yields a fundamentlaly
different model than sequential residual minimization in time. In particular, space--time
residual minimization associates with an algebraic system of equations defined
over all space and time, whose solution comprises a \text{fixed} vector of
(space--time) generalized coordinates; on the other hand, sequential residual
minimization in time (often) associates with a system of ODEs whose solution
comprises a \textit{time-dependent} vector of (spatial) generalized
coordinates.
As a result, space--time residual minimization yields models that are equipped with \textit{a priori} error bounds that grow more slowly
in time, and the resulting \textit{trajectory} exhibits a notion of
optimality; this implies that the solution exhibits a notion of optimality
over the entire time domain. 
\EP{This has nothing to do the basis, which was the point I was trying to get across}
However, these techniques also suffer from several
limitations. First, the computational cost of solving the algebraic system arising from space--time
approaches scales cubically with the number of space--time degrees of freedom;
in contrast, the computational cost incurred by standard
spatial-projection-based ROMs is linear in the number of temporal degrees of
freedom, as the attendant solvers can leverage the lower-block triangular
structure of the system arising from the sequential nature of time
evolution. As a result, solving the algebraic systems arising from space--time
projection is generally intractable without applying hyper-reduction in time
\cite{choi_stlspg,constantine_strom}. Second, space--time residual
minimization precludes future
state prediction, as these methods employ space--time basis vectors defined over the
entire time domain of interest, which must have been included in the training
simulations.

The objectives of this work are to overcome the shortcomings of existing
residual-minimizing methods, and to provide a unifying framework from which
existing methods can be assessed. In essence, the proposed  \textit{\methodNameLower\
(\methodAcronym)} method sequentially minimizes the FOM
residual over a sequence of arbitrarily defined
time windows. The method is characterized by three notable aspects.  First, the method
minimizes the \textit{time-continuous}  residual (i.e., that associated with
the full-order model ODE). By adopting a time-continuous viewpoint, the
formulation decouples the underlying temporal discretization scheme from the
residual-minimization problem, thus addressing a key deficiency of both LSPG
and space--time LSPG. Critically, time-continuous residual minimization also exposes
two different solution methods:  a \textit{discretize-then-optimize}
(i.e., direct) method, and an \textit{optimize-then-discretize} (i.e.,
indirect) method. Second, the method sequentially minimizes
the residual over arbitrarily defined \textit{time windows} rather than
sequentially minimizing the residual over time steps (as in LSPG)
or over the entire time domain (as in space--time residual-minimization
methods). This equips the method with additionally flexibility that enables it
to explore more fine-grained tradeoffs between computational cost and error.
Finally, \methodAcronym\ is formulated for two kinds of space--time trial
subspaces: one that reduces only the spatial dimension of the full-order
model (as employed by traditional spatial-projection-based methods), and one that reduces both the spatial and temporal dimensions of the
full-order model (as employed by space--time methods).
The above attributes allow that \methodAcronym\ method to be viewed as a
generalization of existing model-reduction methods, as 
Galerkin, LSPG, and space--time LSPG correspond to specific
instances of the formulation.
Figure~\ref{fig:flowchart} depicts how the proposed \methodAcronym\ method
provides a unifying framework from which these existing approaches can be derived.

The \methodAcronym\ approach can be viewed as a hybrid space--time method and displays 
commonalities with several related efforts. 
First, Ref.~\cite{bui_thesis} briefly formulated a space--time least-squares ROM
and connected this formulation with optimal control; specifically it mentioned the
optimize-then-discretize vs.\ discretize-then-optimize approaches. This work
did not fully develop this approach, eschewing it 
for 
sequential residual minimization in time (i.e., LSPG). The present work thus
formally develops and extends several of the concepts put forth
in Ref.~\cite{bui_thesis}. Next, Ref.~\cite{constantine_strom} developed a
space--time residual minimization formulation for model interpolation.  The
present work distinguishes itself from Ref.~\cite{constantine_strom} in that
(1) this work also considers trial subspaces characterized by a reduction of
the spatial dimension of the FOM only, and (2) we
minimize the time-continuous FOM residual over arbitrary time
windows. We note that,
similar to the current work, Ref.~\cite{constantine_strom} employs
minimization of the time-continuous FOM residual as a starting point; as such,
this work shares some thematic similarities with
Ref.~\cite{constantine_strom}.  Lastly, Ref.~\cite{choi_stlspg} develops a
space--time extension of the LSPG approach that minimizes the time-discrete
FOM residual over the entire time domain.  The present work distinguishes
itself from Ref.~\cite{choi_stlspg} in that (1) this work minimizes the
time-continuous FOM residual, (2) this work minimizes this residual over
arbitrary time windows, and (3) this work also considers trial subspaces
characterized by a reduction of the spatial dimension of the FOM only.

@@

 In summary, specific contributions of this work include:
\begin{enumerate}
%\item The development of a time-continuous residual minimization framework for reduced-order models.
\item The \methodNameLower\ (\methodAcronym) framework for dynamical-system
	model reduction. The approach sequentially minimizes the time-continuous
		full-order model residual over arbitrary time windows.
\item Support of two space--time trial subspaces: one that reduces the spatial
	dimension of the FOM only, and one that reduces both the spatial and
		temporal dimensions of the FOM. The former case is of particular
		interest in the \methodAcronym\  contex, as the stationary conditions are
		derived via the Euler--Lagrange equations and comprise a coupled two-point
		Hamiltonian boundary value problem containing a forward and backward
		system. The forward system, which is forced by an auxiliary costate,
		evolves the (spatial) generalized coordinates of the ROM in time. The
		backward system, which is forced by the time-continuous FOM residual
		evaluated about the ROM state, governs the dynamics of the costate. 
	\item Derivation of two solution techniques:
		\textit{discretize-then-optimize} (i.e., direct) and
		\textit{optimize-then-discretize} (i.e., indirect). 
	\item Derivations of conditions under which the
		\methodAcronym\ framework recovers Galerkin, LSPG, and space--time LSPG.
\item Error analysis of the \methodAcronym\ framework using
	trial subspaces characterized by a reduction of
the spatial dimension of the FOM only.
	This analysis demonstrates that, over a given window, the \methodAcronymROMs\ error 
		is bounded \textit{a priori} by a combination of the
		error at the start of the window and the integrated FOM ODE residual
		evaluated at the FOM state projected onto the trial subspace. 
\EP{Updated this, I don't show what you said}%As a result,
%		the error bound exhibits exponential growth in the number of time windows
%		(not total time instances), yielding smaller error-bound growth in thime than
%		Galerkin or LSPG projection. \KTC{Check this}
\item Numerical experiments for 
	trial subspaces characterized by a reduction of
the spatial dimension of the FOM only, which demonstrate two key findings:
\begin{itemize}
\item Minimizing the residual over a larger time window leads to more stable solutions 
with lower space--time residuals norms.
\item Minimizing the residual over a larger time window does not necessarily
	lead to a more accurate trajectory (as measured in the space--time
		$\elltwo$-norm of the solution). Instead, minimizing the residual over an
		intermediate-sized time window leads to the smallest trajectory error.
\end{itemize}
\end{enumerate}

The paper proceeds as follows: Section~\ref{sec:math}
outlines the mathematical setting for the full-order model, along with the Galerkin, LSPG, and
space--time LSPG approaches. Section~\ref{sec:tclspg} outlines the proposed \methodAcronym\
framework. Section~\ref{sec:numerical_techniques}
outlines numerical techniques for solving \methodAcronymROMs, including both direct and
indirect methods. Section~\ref{sec:analysis} provides
equivalence conditions and error analysis for \methodAcronymROMs.
Section~\ref{sec:numerical_experiments} presents numerical experiments.
Section~\ref{sec:conclude} provides conclusions and perspectives.
We denote vector-valued functions with italicized bold symbols (e.g., $\boldsymbol
x$), vectors with standard bold symbols (e.g., $\mathbf{x}$), 
matrices with capital bold symbols (e.g., $\mathbf{X} \equiv \begin{bmatrix}
\mathbf{x}_1 & \cdots & \mathbf{x}_r\end{bmatrix}$), and spaces with
calligraphic symbols (e.g., $\mathcal{X}$). We additionally denote differentiation of a time-dependent 
function with respect to with the $\dot\ $ operator.
%Model reduction of nonlinear dynamical systems is a growing research area.
%The least-squares Petrov-Galerkin (LSPG) approach has gained popularity as a model reduction technique for nonlinear dynamical systems. Derived at the fully discrete level, the LSPG approach solves a least-squares problem that minimizes the fully discrete residual of the reduced-order model at each time-step. LSPG has been shown to be an effective formulation for nonlinear model reduction on complex problems of interest.
%\begin{comment} 
\section{Mathematical formulation}\label{sec:math}
	We begin by providing the formulation for the full-order model,
	followed by a description of standard model-reduction methods
	classified according to the type of trial subspace they employ.
	\subsection{Full-order model}\label{sec;FOM}
We consider the full-order model to be a dynamical system expressed as a
	system of ordinary differential equations (ODEs)
\begin{equation}\label{eq:FOM}
 \stateFOMDotArg{t}  = \velocity(\stateFOMArg{}{t},t), \qquad \stateFOM(0) =
	\stateFOMIC,\qquad t \in [0,T]
\end{equation}
where  $\stateFOM: [0,T]
	\rightarrow  \RR{\fomdim}$ with $\stateFOM: \timeDummy \mapsto
	\stateFOM(\timeDummy) $ and $\stateFOMDot \equiv {d \stateFOM}/{d\tau}$
	denotes the state implicitly defined as the solution to initial value
	problem~\eqref{eq:FOM}, 
$T \in \mathbb{R}^+$ denotes the final time, 
 $\stateFOMIC \in \mathbb{R}^{\fomdim}$ denotes the initial condition,
	and $\velocity: \mathbb{R}^{\fomdim} \times [0,T] \rightarrow
	\mathbb{R}^{\fomdim}$ with $(\stateyDiscrete,\timeDummy) \mapsto
	\velocity(\stateyDiscrete,\timeDummy)$ denotes the velocity, which is possibly
	nonlinear in its first argument. For subsequent exposition, we introduce
	$\timeSpace$ to denote the set of real-valued functions acting on the time
	domain (i.e., $\timeSpace = \{f\,|\,f:[0,T]\rightarrow\RR{}\}$); the
	state can be expressed equivalently as $\stateFOM \in \RR{\fomdim} \otimes
	\timeSpace $.  We refer to the initial value problem defined in
	Eq.~\eqref{eq:FOM} as the ``full-order model" (FOM) ODE. We note that
	although the problem of interest described in the introduction corresponds
	to a parameterized dynamical system, we suppress dependence of the FOM ODE
	\eqref{eq:FOM} on such parameters for notational convenience, as this work
	focuses on devising a model-reduction approach applicable to a specific parameter instance.
